% ---------------------------------------------------------
% Project: PhD KAPPA
% File: acknowledgements.tex 
% Author: Andrea Discacciati
%
% Purpose: Acknowledgements
% ---------------------------------------------------------

\chapter{Acknowledgements}

There are many people that I would like to thank for their contributions to this thesis, and for their support and encouragement during these  years.

\bigskip

\textbf{Nicola Orsini}, my main supervisor for the second half of my doctoral education. Thank you for everything you have taught me, for helping me put things into perspective, and for your genuine and contagious enthusiasm. You have been not only an outstanding supervisor, but also a sincere friend. For this and a lot more, I owe a great debt of gratitude to you.

\bigskip

\textbf{Alicja Wolk}, my main supervisor for the first half of my PhD studies. Thank you for taking me on as a PhD student, for teaching me about the world of research and epidemiology, and for always giving me the freedom to follow my interests. Your guidance has been invaluable.

\bigskip

My co-authors \textbf{Matteo Bottai} and \textbf{Sander Greenland} for everything you have taught me about statistics. \textbf{Andrea Bellavia} and  \textbf{Alessio Crippa} for constructive discussions. \textbf{Swen-Olof Andersson}, \textbf{Ove Andrén}, and \textbf{Jan-Erik Johansson} for sharing your knowledge on prostate cancer. To all of you, I am grateful for the invaluable contributions you made to the studies included in this thesis.

\bigskip

My colleagues, past and present, at the Unit of Nutritional Epidemiology.  \textbf{Bettina Julin} and \textbf{Laura Thomas}, who made me feel at home from the very first day. \textbf{Daniela Di~Giuseppe}, who let me know about the possibility of pursuing a PhD in this unit. \textbf{Andrea Bellavia}, \textbf{Alessio Crippa}, and \textbf{Viktor Oskarsson}, with whom I have shared office, ideas, thoughts, and laughs.  An extra warm thank you goes to \textbf{Charlotte Bergkvist},  \textbf{Niclas Håkansson}, \textbf{Susanne Rautiainen Lagerström}, and \textbf{Alice Wallin}. Thank you to \textbf{Katica Anusic}, \textbf{Ann Burgaz}, \textbf{Holly Harris}, \textbf{Jennifer Höglund}, \textbf{Anna Ingemarsdotter}, \textbf{Heather Jonasson}, \textbf{Susanna Larsson}, \textbf{Becky Leung}, \textbf{Nalleli Vivanco Muñoz},  \textbf{Camilla Olofsson}, \textbf{Emma Patterson}, \textbf{Iffat Rahman}, \textbf{Jinjin Zheng Selin}, \textbf{Lollo Sjöholm}, \textbf{Otto Stackelberg}, \textbf{Frej Stilling}, \textbf{Thanasis Tektonidis}, \textbf{Mimi Throne-Holst}, \textbf{Sara Wallenius}, and \textbf{Agneta Åkesson} (plus everyone else whom I may have forgotten to mention here) for making the office such a pleasant place to be.

\bigskip

My colleagues at the Unit of Biostatistics: \textbf{Davide Bossoli}, \textbf{Silvia Columbu}, \textbf{Xin Fang}, \textbf{Daniel Olsson}, \textbf{Celia Garcia Pareja}, and \textbf{Michele Santacatterina}. In particular, I would like to thank \textbf{Paolo Frumento} for always taking the time to answer my questions.

\bigskip

Participants and faculty of the Summer School on Modern Methods in Biostatistics and Epidemiology. In particular, I would like to thank \textbf{Marcello Pagano}, \textbf{Rino Bellocco}, \textbf{Marco Bonetti}, and \textbf{Nicola Orsini} for the opportunity to take part in the summer school and for the privilege to serve as teaching assistant for some of the courses. It has been a terrific experience.

\bigskip

\textbf{Anita Berglund}, for the organization of courses and workshops in beautiful Åre. Thank you for giving me the chance to participate. 

\bigskip

My parents, \textbf{Edoardo} and \textbf{Anna}, for the support and chances you have given me over the years.  

\bigskip

Above all, I would like to thank \textbf{Francesca}. Words cannot express my gratitude for your unbounded love, infinite patience, and unremitting encouragement. 

\vfill

\noindent This work was supported by \textbf{Karolinska Institutet}'s funding for doctoral students (\mbox{KID-funding}), \textbf{Vetenskapsrådet}/Rådet för forskningens infrastrukturer (the Swedish Research Council/Committee for Infrastructure), and \textbf{Cancerfonden} (the Swedish Cancer Society).






